
```{r}
# install.packages("anytime")
# install.packages("plyr")
# installed.packages("anytime")
# install.packages("readxl")
# install.packages("xlsx")
library(MASS)
library(anytime)
library(tidyverse)
library(lubridate)
library(dplyr)
library(readxl)
library(ggplot2)
library(gridExtra)
library(plyr)
library(ggpubr)
```


```{r}
data <- data.table::fread("/Users/oliviashi/Documents/MASTER_RESEARCH/data/ca.csv")
data$date <- as.Date(data$date,'%d-%m-%Y')
d <- data %>% filter(prname=="Quebec")
s3 <- read_excel("/Users/oliviashi/Documents/MASTER_RESEARCH/data/covid-19-intervention-scan-data-tables-en.xlsx",sheet=3,col_names=TRUE,skip=2)
s3$date <- as.Date(s3$`Start date`)
s3$cat <- s3$`Intervention Category`
data <- s3 %>% dplyr::select("Entry ID","date","cat","Action","Jurisdiction","Description") %>% filter(Jurisdiction=="Que.")
data <- tibble(data)
data$Closures <- as.numeric(NA)
data$Openings <- as.numeric(NA)
s3 %>% filter(cat=="Closures/openings") %>% filter(Jurisdiction=="Que.") %>% filter(Action=="Eased")

#join tables
da<-join(d,data,by="date",type="left",match="first")
i=1
for (i in 1:dim(da)[1]){
if (isTRUE(da$Action[i] == "New") && isTRUE(da$cat[i]=="Closures/openings")) {
da$Closures[i] <- da$cases[i]}
else if (isTRUE(data$Action[i] == "Eased") && isTRUE(da$cat[i]=="Closures/openings")){
da$Openings[i] <- da$cases[i]}}

# ggplot(data=da,aes(x=date))+geom_line(aes(x=date,y=casesMA),col="blue") +  geom_point(aes(y=as.numeric(`Closures`),col="Closures")) + geom_point(aes(y=as.numeric(`Openings`),col="Openings")) + ggtitle(sprintf("cases and intervention plot for %s",d$prname))
```
else{
    data$`Closures`[i]<-"other"
  }
```{r}
data <- da %>% dplyr::select(date,casesMA,cat,Action,Description)
data <- pivot_wider(data=data,names_from=cat,values_from=Description) 
data$Closures<-NA
data$Openings<-NA
data$cases2<-NA
for(i in 1:length(data$date)){
  if(is.na(data$`Closures/openings`[i])){
    data$`Closures/openings`[i]<-NA
  }else if(grepl('school',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="New"){
    data$Closures[i]<-"school closed"
  }else if(grepl('school',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="Eased"){
    data$Openings[i]<-"school opened"
  }else if(grepl('restaurant',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="New"){
    data$Closures[i]<-"restaurant, rec closed"
  }else if(grepl('restaurant',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="Eased"){
    data$Openings[i]<-"restaurant, rec opened"
  }else if(grepl('recreational',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="New"){
    data$Closures[i]<-"restaurant, rec closed"
  }else if(grepl('recreational',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE) && da$Action[i]=="Eased"){
    data$Openings[i]<-"restaurant, rec opened"
  }else if(grepl('non-essential',data$`Closures/openings`[i], ignore.case = TRUE) && da$Action[i]=="New"){
    data$Closures[i]<-"non-essential closed"
  }else if(grepl('non-essential',data$`Closures/openings`[i], ignore.case = TRUE) && da$Action[i]=="Eased"){
    data$Openings[i]<-"non-essential opened"
  }else if((grepl('public places',data$`Closures/openings`[i], ignore.case = TRUE) ||grepl('parks',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE)||grepl('museums',data$`Closures/openings`[i], ignore.case = TRUE)) && da$Action[i]=="New"){
    data$Closures[i]<-"public places closed"
  }else if((grepl('public places',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE)||grepl('parks',data$`Closures/openings`[i], ignore.case = TRUE, fixed=FALSE)||grepl('museums',data$`Closures/openings`[i], ignore.case = TRUE)) && da$Action[i]=="Eased"){
    data$Openings[i]<-"public places opened"
  }
}

for(i in 1:length(data$date)){
  if(!is.na(data$`Closures/openings`[i])){
    data$cases2[i] <- data$casesMA[i]
  }
}

data$cases3<-NA

for(i in 1:length(data$date)){
  if(!is.na(data$Distancing[i])){
    data$cases3[i] <- data$casesMA[i]
  }
}

for(i in 1:length(data$date)){
  if(is.na(data$Distancing[i])){
  }else if(grepl('curfew',data$Distancing[i], ignore.case = TRUE) && da$Action[i]=="New"){
    data$Distancing[i]<-"curfew"
  }else if(grepl('gathering',data$Distancing[i], ignore.case = TRUE) && da$Action[i]=="New"){
    data$Distancing[i]<-"gathering restricted"
  }else if(grepl('gathering',data$Distancing[i], ignore.case = TRUE) && da$Action[i]=="Eased"){
    data$Distancing[i]<-"gathering permitted"
  }else if(grepl('worship',data$Distancing[i], ignore.case = TRUE) && da$Action[i]=="New"){
    data$Distancing[i]<-"bar, worship restricted"
  }else if(grepl('bar',data$Distancing[i], ignore.case = TRUE) && da$Action[i]=="New"){
    data$Distancing[i]<-"bar, worship restricted"
  }else{
    data$Distancing[i]<-"other"
  }
}
```

fig1 <- ggplot(data=data,aes(x=date))+geom_line(aes(y=casesMA))+geom_point(aes(y=cases2,group=`Closures/openings`,colour=Closures,alpha=Openings))+ scale_alpha_discrete(range = c(.5, 1),na.translate=F) +scale_x_date(date_labels="%m-%y",date_breaks="2 month")+ ggtitle("Closures and Openings for Quebec") + scale_colour_discrete(na.translate = F)


```{r}
fig1 <- ggplot(data=data,aes(x=date))+geom_line(aes(y=casesMA))+geom_point(aes(y=cases2,group=`Closures/openings`,colour=Closures)) +scale_x_date(date_labels="%m-%y",date_breaks="2 month")+ ggtitle("Closures and Openings for Quebec") + scale_colour_discrete(na.translate = F)

fig2 <- ggplot(data=data,aes(x=date))+geom_line(aes(y=casesMA))+scale_x_date(date_labels="%m-%y",date_breaks="2 month")+geom_point(aes(y=cases3,colour=Distancing)) + ggtitle("Distancing for Quebec")+  scale_colour_discrete(na.translate = F)

ggarrange(fig2,fig1,nrow=2)
fig1
```
```{r}
data
```























