```{r message=FALSE, warning=FALSE}
library(lubridate)
library(dplyr)
library(EpiEstim)
library(tidyverse)
library(zoo)
library(ggplot2)
library(reshape2)
library(ggpubr)
cases<-read.csv('/Users/oliviashi/Documents/MASTER_RESEARCH/Covid19Canada/timeseries_prov/cases_timeseries_prov.csv')%>% dplyr::select(province,date_report,cases) %>% filter(province=='Quebec')
names(cases)[2]<-"date"
testing<-read.csv('/Users/oliviashi/Documents/MASTER_RESEARCH/Covid19Canada/timeseries_prov/testing_timeseries_prov.csv')  %>% filter(province=='Quebec') %>% dplyr::select(date_testing,testing,cumulative_testing)
names(testing)[1]<-"date"

data=merge(x=cases,y=testing,by="date")
data$date <- as.Date(data$date,format="%d-%m-%Y", tryFormats= "%Y-%m-%d")
for(i in 1:dim(data)[1]){
  if(data$testing[i]==0){
    data$testing[i]=data$testing[i-1]
  }}
data$cases[1]=10000
for(i in 1:dim(data)[1]){
  if(data$cases[i]<=0){
    data$cases[i]=data$cases[i-1]
  }}

findPositivityRate <- function(x,y){
  res <- 0
  for(i in 1:length(x)){res[i] <- ifelse(x[i]<y[i],x[i]/y[i],1)}
  return(res)
}
data$positivity <- findPositivityRate(data$cases,data$testing)

data<-data %>% arrange(date)
data$cases<-rollmean(data$cases,14,align="left",fill=1000)
data$testing<-rollmean(data$testing,14,align="left",fill=10000)
data$cases<-rollmean(data$cases,2,align="left",fill=1000)
data$testing<-rollmean(data$testing,7,align="left",fill=10000)

findDayi <- function(x){
  res<-0
  for(i in 1:length(x)){res<-difftime(x,as.Date("2020-03-15"),units="days")}
  return(as.numeric(res))
}
data$dayi<-findDayi(data$date)

for(i in 1:length(data$dayi)){
  data$prevalence[i] <- (1500 / (data$dayi[i] + 50)) * (data$positivity[i])^(0.5) + 2
  data$a[i] <- 1500 / (data$dayi[i] + 50)}

data$trueNewDailyInfections <- data$cases * data$prevalence
for(i in 2:dim(data)[1]){
  if(is.nan(data$trueNewDailyInfections[i])|data$trueNewDailyInfections[i]==0.0000){
    data$trueNewDailyInfections[i]=data$trueNewDailyInfections[i-1]}}
for(i in 10:dim(data)[1]){
  if(data$trueNewDailyInfections[i]>data$trueNewDailyInfections[i-1]+data$trueNewDailyInfections[i-2]+data$trueNewDailyInfections[i-3]*0.5){
    data$trueNewDailyInfections[i]=data$trueNewDailyInfections[i-1]*1.4
    data$trueNewDailyInfections[i-1]=data$trueNewDailyInfections[i-2]*1.4
    data$trueNewDailyInfections[i-2]=data$trueNewDailyInfections[i-3]*1.4}
}
```


fig 2.2
```{r}
data<-data%>% filter(date<'2021-12-01')%>% filter(date>'2020-04-01')
fig1<-ggplot(data=data,aes(date)) +
  geom_line(aes(y = cases, colour = "Reported cases")) + 
  geom_line(aes(y = trueNewDailyInfections, colour = "Actual cases"))+
  scale_x_date(breaks = "2 months", date_labels = "%b/%y")+
theme(legend.background=element_rect(fill = alpha("white", 0.5)),legend.title = element_blank(),legend.position = c(0.85, 0.8),legend.text = element_text(size = 10),legend.key.size=unit(0.5,"lines"),legend.key = element_blank(),axis.title.x=element_blank())
  
plt3<-ggplot(data=data,aes(x=date))+
  geom_line(aes(y=cases,col="Reported cases"))+
  geom_line(aes(y=testing,col="Total tests"))+
  ylim(0,20000)+
  scale_x_date(breaks = "2 months", date_labels = "%b/%y")+
theme(legend.background=element_rect(fill = alpha("white", 0.5)),legend.title = element_blank(),legend.position = c(0.85, 0.8),legend.text = element_text(size = 10),legend.key.size=unit(0.5,"lines"),legend.key = element_blank(),axis.title.x=element_blank())
#ggtitle("Reported cases vs. total tests")+

actual_vs_reported<-ggarrange(plt3, fig1, nrow = 2,labels = c("A","B")) 
actual_vs_reported
ggsave("../LatexImages/actual_vs_reported.png",actual_vs_reported,scale=1)
knitr::plot_crop("../LatexImages/actual_vs_reported.png")
```


g<-ggplot(data=df_figure,aes(x=date)) + geom_line(aes(y=casesSmooth,col="Reported cases"))+  geom_line(aes(y=casesPRr,col="Infections (Prevalence Ratio)"))+ geom_line(aes(y=casesRC,col="Infections (Richard-Lucy)"))+
  ggtitle("Comparisons of True New Daily Infections from two different methods")+
  scale_x_date(breaks = "2 months", date_labels = "%b %y")+ theme(legend.position = c(0.85, 0.8),legend.title = element_text(size = 7),legend.text = element_text(size = 7),legend.key.size=unit(0.5,"lines"),legend.key = element_blank())
ggarrange(g,g,nrow=2)






```{r}
# plt1<-ggplot(data=data,aes(x=dayi,a)) + geom_point()+
#       ggtitle("Effect on day_i on exponential constant a in Quebec")+
#   scale_x_date(breaks = "2 months", date_labels = "%b %y")+
#   theme(legend.position = c(0.8, 0.8),legend.title = element_text(size = 5),legend.text = 
# 	element_text(size = 5),legend.key.size=unit(0.5,"lines"),legend.key = element_blank())

# plt2<-ggplot(data=data,aes(x=positivity,y=prevalence))+geom_point()+
#       ggtitle("Relationship between test positivity and prevalence ratio in Quebec")

# plt3<-ggplot(data=data,aes(x=positivity,y=prevalence))+geom_point()+ggtitle("Relationship between test positivity and prevalence ratio in Quebec")

```


```{r}
plt3<-ggplot(data=data,aes(date)) +
      geom_line(aes(y = cases, colour = "Reported Infections")) + 
      geom_line(aes(y = trueNewDailyInfections, colour = "Adjusted Infections")) +
      geom_line(aes(y = testing, colour = "Total Tests ")) +
      ggtitle(sprintf("Reported vs. adjusted cases in Quebec"))
plt3
```






```{r}


plt2 <- ggplot(data=data,aes(x=date,y=positivity)) + geom_point() #+ ggtitle("positivity rate in ",Quebec$prname[1])

plt3<-ggplot(data=qc,aes(x=positivity,y=prevalence))+geom_point()


ggarrange(plt1, plt2, labels = c("A","B"), ncol = 1, nrow = 2) 
```




















































